Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-16 paper.qmd
Updated Figure 15
Updated Figure 10

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is universally poor. Layouts a that show two close clusters are universally suboptimal. Layout a is the best choice."
#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
#| fig-pos: H
#| out-width: 60%

@@

```{r}
#| label: fig-one-dens_clust-error
#| fig-height: 10
#| fig-width: 10
#| fig-pos: H
#| out-height: 50%
#| fig-cap: "The sparse end of the structure exhibits a high model error, while the dense end shows a low model error. The tSNE layouts of the C-shaped structure, with uniform points (a1) and dense points (b1) are colored according to their high-dimensional error, along with their selected projections (a2 and b2), are displayed. Darker colors represent a high error, while lighter colors indicate a low error. The uniformly distributed points demonstrate random model errors throughout the structure, whereas the dense structure shows large model error at the sparse end. Videos of the langevitour animations are available at."

@@

To showcase the robustness of our method when applied to data with different characteristics, this section presents two case studies. The first application focuses on single-cell data, a domain where the high dimensionality, sparsity, and heterogeneity of gene expression measurements pose significant analytical challenges. Our approach facilitates the identification of the most accurate NLDR layout, enabling the discovery of cell populations with similar expression profiles. The second case study explores the MNIST hand-written digits dataset, which serves as a benchmark for evaluating the preservation of local structures, specifically within a selected digit.

@@

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the MNIST digit 1 data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). All the layouts appear to be very similar. Layout c is universally poor. Layouts f that show two close clusters are universally suboptimal. Layout f is the best choice."
#| label: fig-mnist-mse
#| fig-pos: H
#| out-width: 80%

>>>>

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is universally poor. Layouts a, b, e that show two close clusters are universally suboptimal. Layout b with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. Layout e has small separation with oddly shaped clusters. Layout a is the best choice."
#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
#| fig-pos: H
#| out-width: 60%

@@

```{r}
#| label: fig-one-dens_clust-error
#| fig-height: 10
#| fig-width: 10
#| fig-pos: H
#| out-height: 50%
#| fig-cap: "The sparse end of the structure exhibits a high model error, while the dense end shows a low model error. The tSNE layouts of the C-shaped structure, with uniform points (a1) and dense points (b1) are colored according to their high-dimensional error, along with their selected projections (a2 and b2), are displayed. Darker colors represent a high error, while lighter colors indicate a low error. The uniformly distributed points demonstrate random model errors throughout the structure, whereas the dense structure shows large model error at the sparse end. Videos of the langevitour animations are available at <>."

@@

To showcase the robustness of our method when applied to data with different characteristics, this section presents two case studies. The first application focuses on single-cell data, a domain where the high dimensionality, sparsity, and heterogeneity of gene expression measurements pose significant analytical challenges. Our approach facilitates the identification of the most accurate NLDR layout, enabling the discovery of cell populations with similar expression profiles. The second case study explores the MNIST hand-written digits dataset, which serves as a benchmark for evaluating the preservation of local structures, specifically with digit 1.

@@

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the MNIST digit 1 data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). All the layouts appear to be very similar. Layout c is universally poor. Layouts a, f that show two close clusters are universally suboptimal. Layout a with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. Layout f is the best choice."
#| label: fig-mnist-mse
#| fig-pos: H
#| out-width: 80%
